Optimal design of experiment

### 2024-07-09 Sections/Optimal_design_of_experiment.tex
09.07

	Details on the process model can be found in $({\color{red}article~1})$. The model's empirical correlations are derived from laboratory experiments conducted under various constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. This study employs a model-based design of experiments technique to design an experiment with dynamically changing conditions to improve the precision of the correlation for $D_i$. The decision variables are adjusted every 30 minutes and are kept constant within each interval (piecewise constant controls). These controls have lower and upper bounds that match the validated range of the process model detailed in $({\color{red}article~1})$. The sampling time is considered to be 10 minutes, and the total extraction time is 300 minutes. The standard deviation $\sigma^2$ was estimated to be 0.12. The analysed correlations consist of two independent variables: Reynolds number and mass flow rate. The Reynolds number is a function of fluid thermodynamic properties and velocity, and the fluid density can be affected by temperature or pressure changes. In this work, two cases are solved and compared against each other.

	The first case assumes that the decision variables are the inlet temperature ($T^{in}$) and mass flow rate ($F$). Throughout the experiment, the pressure is assumed to remain constant and is set to a value in the middle of the validated range ($150$ bar). The initial state is considered isothermal, so $T^0=T^{in}(t=0)$. The general formulation of the optimal design of the experiment problem is presented below.

@@

	The second case assumes that the pressure and mass flow rate are treated as decision variables while the inlet temperature is kept constant at $35^\circ$C. Similarly to the first case, the extractor is assumed to be isothermal at the beginning of the operation. Compared to the previous case, the pressure change affects the temperature inside of the extractor. Any deviation in extractor temperature caused by pressure change will eventually be cancelled out by the inflow of fresh fluid at the designed temperature of $35^\circ$C. The general formulation of the optimal design of the experiment problem is presented below.

>>>>

	Details on the process model can be found in [article 1]. The model's empirical correlations are derived based on laboratory experiments conducted under various constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. This study employs a model-based design of experiments technique to design an experiment with dynamically changing conditions to improve the precision of the correlation for $D_i$. The decision variables are adjusted every 10 minutes and are kept constant within each interval (piecewise constant controls). These controls have lower and upper bounds that match the validated range of the process model detailed in [article 1]. The sampling time is set to 5 minutes, and the total extraction time is 300 minutes. The standard deviation $\sigma^2$ was estimated to be 0.12. The analysed correlations consist of two independent variables: Reynolds number and mass flow rate. The Reynolds number is a function of fluid thermodynamic properties and velocity, and the fluid density can be affected by temperature or pressure changes. In this work, both cases are solved and compared.

	In the first case, the decision variables are the inlet temperature $(T^{in})$ and mass flow rate $(F)$. Throughout the experiment, the pressure is assumed to remain constant at 150 bar (the middle of the validated range). The initial state is considered isothermal, so $T^0 = T^{in}(t=0)$. The problem formulation for Case 1 is given by Equation \ref{EQ:Formulation_1}.

@@

	In the second case, the pressure and mass flow rate are treated as decision variables, while the inlet temperature is kept constant at $35^\circ C$. Similar to the first case, the extractor is assumed to be isothermal at the beginning of the operation. Unlike the previous case, changes in pressure affect the temperature inside the extractor. Any temperature deviation caused by pressure changes will eventually be mitigated by the inflow of fresh fluid at the constant temperature of $35^\circ C$. The problem formulation for Case 2 is given by Equation \ref{EQ:Formulation_2}.

### 2024-07-09 Sections/Results_DOE.tex
09.07

	%To identify the global solution of the Equation \ref{EQ:Formulation}, the optimization problem is solved multiple times starting from random initial solutions. The solution with the lowest value of the cost function is considered to be the global solution. Figure \ref{fig:scatter}, show an compression between initial and the final values of the cost function for multiple optimization problems. The shaded area is spanned between the smallest and the third smallest final value of the objective function. The black curves on the left and below of the scatter plot indicate distributions of the initial and final values of the cost function, respectively.

	To identify the global solution of Equation \ref{EQ:Formulation}, the optimization problem is solved multiple times starting from random initial solutions. The solution with the lowest cost function value is considered the global solution. Figure \ref{fig:scatter} shows a comparison between the initial and final values of the cost function for multiple optimization runs. The shaded area spans from the smallest to the third smallest final value of the objective function. The black curves on the left and below the scatter plot indicate the distributions of the initial and final cost function values, respectively.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=\columnwidth]{Figures/Results/scatter.png}	
		\caption{Initial vs final values of the cost function}
		\label{fig:scatter}
	\end{figure}

	The optimal inlet temperature and mass flow rate profile are presented on Figure \ref{fig:profile}.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=\columnwidth]{Figures/Results/Profile.png}	
		\caption{Optimal temperature and mass flow rate profiles}
		\label{fig:profile}
	\end{figure}

	\begin{comment}
		\begin{figure}[h!]
			\centering
			\includegraphics[width=\columnwidth]{Figures/Results/Profile_2.png}	
			\caption{2nd Optimal temperature and mass flow rate profiles}
			\label{fig:profile_2}
		\end{figure}

		\begin{figure}[h!]
			\centering
			\includegraphics[width=\columnwidth]{Figures/Results/Profile_3.png}	
			\caption{3rd Optimal temperature and mass flow rate profiles}
			\label{fig:profile_3}
		\end{figure}
	\end{comment}

>>>>

	To identify the global solution of Equations \ref{EQ:Formulation_1}, the optimization problem is solved multiple times, each starting from a random initial solution. Figure \ref{fig:scatter_T} compares the initial and final values of the cost function across multiple optimization runs. The shaded area spans from the smallest to the third smallest final values of the objective function. The black curves on the left and below the scatter plot represent the distributions of the initial and final cost function values, respectively. From Figure \ref{fig:scatter_T}, it can be observed that multiple local solutions exist for Case 1. The solution with the lowest cost function value is considered the global solution.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=\columnwidth]{Figures/Results/scatter_T.png}	
		\caption{Initial vs final values of the cost function}
		\label{fig:scatter_T}
	\end{figure}

	The optimal profiles of inlet temperature and mass flow rate are presented in Figure \ref{fig:profile_T}. One of the independent parameters in the analysed correlation is the mass flow rate, which is directly modified several times during the experiment. The second independent parameter is the Reynolds number, which can be affected by changes in the thermodynamic properties of the fluid. The thermal properties of CO$_2$ between $30^\circ$C and $40^\circ$C, assuming a constant pressure of 150 bar, vary only slightly. For instance, the dynamic viscosity changes from 67 to $80 \times 10^{-6}$ Pa$\cdot$s. The obtained profile of the inlet temperature shows minimal variation, suggesting a limited influence of inlet temperature on the extraction yield. Given the narrow operational window for inlet temperature at constant pressure, the mass flow rate becomes the primary variable to control.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=\columnwidth]{Figures/Results/Profile_T_1.png}	
		\caption{Optimal temperature and mass flow rate profiles}
		\label{fig:profile_T}
	\end{figure}

	Similar to Case 1, Case 2 is solved multiple times starting from random initial solutions to identify the global solution. Figure \ref{fig:scatter_P} summarizes the changes in the objective function values across all optimization runs. It is evident that both the initial and final values of the cost function are significantly lower than any values obtained in Case 1. In Case 2, the wider operating range leads to more significant changes in the physical properties of CO$_2$ compared to Case 1. For example, the kinematic viscosity of CO$_2$ ranges from $48 \cdot 10^{-6}$ (at $40^\circ$C and 100 bar) to $90 \cdot 10^{-6}$ Pa$\cdot$s (at $30^\circ$C and 200 bar). The greater variability in pressure induces a more complex system response, including temperature variations, resulting in lower values of the objective function compared to Case 1.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=\columnwidth]{Figures/Results/scatter_P.png}	
		\caption{Initial vs final values of the cost function}
		\label{fig:scatter_P}
	\end{figure}

	The optimal profiles of pressure and mass flow rate are presented on Figure \ref{fig:profile_P}. Similarly to case 1, only few changes in the mass flow rate are observed in the plot. In contrary to the inlet temperature, the pressure changes are frequently applied in the case 2.

	\begin{figure}[h!]
		\centering
		\includegraphics[width=\columnwidth]{Figures/Results/Profile_P_1.png}	
		\caption{Optimal pressure and mass flow rate profiles}
		\label{fig:profile_P}
	\end{figure}
